Discrete Time Markov Chains.tex

%----------------------------------%

\section{Markov Chains}
A Markov chain is a stochastic process with the Markov property. The term "Markov chain" refers to the sequence of random variables such a process moves through, with the Markov property defining serial dependence only between adjacent periods (as in a "chain"). It can thus be used for describing systems that follow a chain of linked events, where what happens next depends only on the current state of the system.

In the literature, different kinds of Markov process are designated as "Markov chains". Usually the term is reserved for a process with a discrete set of times, i.e. a \textbf{discrete-time Markov chain (DTMC)}. On the other hand, a few authors use the term "Markov process" to refer to a continuous-time Markov chain without explicit mention.

While the time parameter is usually discrete, the state space of a Markov chain does not have any generally agreed-on restrictions: the term may refer to a process on an arbitrary state space.

However, many applications of Markov chains employ finite or countably infinite (that is, discrete) state spaces, which have a more straightforward statistical analysis. Besides time-index and state-space parameters, there are many other variations, extensions and generalisations. 

For simplicity, most of this article concentrates on the discrete-time, discrete state-space case, unless mentioned otherwise.

\subsection{Transitions}
The changes of state of the system are called transitions. The probabilities associated with various state changes are called transition probabilities. The process is characterized by a state space, a transition matrix describing the probabilities of particular transitions, and an initial state (or initial distribution) across the state space. By convention, we assume all possible states and transitions have been included in the definition of the process, so there is always a next state, and the process does not terminate.

%------------------------------------------------------------------%
\section{Expected Hitting Time}

%%---http://www.hamilton.ie/ollie/Downloads/Mar1.pdf
